---
title: "The Languages of RStudio"
subtitle: 'A few of many'
output:
  html_document:
    self_contained: true
    fig_caption: no
    theme: 'spacelab'
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document illustrates the versatility of RStudio by performing the same operation in one document with each of R, Python, Bash, and PowerShell.

Retrieved data is then loaded into SQLite and queried with SQL.

Lastly, see how to producer "pretty" data tables, and also how interoperability between R and Python can easily be achieved with the `reticulate` package.

## R

```{r}
df <- read.csv('users.csv')

df
```

## Python

```{python}
import pandas as pd

df = pd.read_csv('users.csv')

df
```

## Bash

```{bash}
input="users.csv"
while IFS=',' read -r f1 f2 f3
do 
  echo "$f1 $f2 $f3"
done < "$input"
```

[source](https://www.cyberciti.biz/faq/linux-unix-appleosx-bsd-shell-parse-text-file/)

## PowerShell

```{bash, engine.path='/usr/bin/pwsh'}
Import-Csv 'users.csv'
```

## SQL

First, create the SQLite db, then load our dataframe `df` into it.

```{r}
library(RSQLite)

con <- dbConnect(RSQLite::SQLite(), dbname=':memory:')

dest_table <- 'df_table'

dbWriteTable(con, name=dest_table, value=df)
```

Query the new table in a SQL code chunk.

```{sql, connection=con}
SELECT *
FROM ?dest_table
LIMIT 5
```

```{r, echo=FALSE}
dbDisconnect(con)
```

SQL code chunks output a pretty HTML table by default. Keep reading to see how to present your R/Python data frame(s) in a pretty table.

## Pretty Tables

The R package `knitr` provides the function `kable` to produce a HTML table from an R data frame.

Similarly, the R package `DT` provides the function `datatable` as an interface to the popular [DataTables JavaScript library](https://datatables.net/), which presents data frames in an interactive HTML widget.

```{r}
library(knitr) # for kable
library(DT)    # for datatable
```

### Kable

*R*

```{r}
knitr::kable(df)
```

*Python*

> Simply load the `reticulate` package to achieve interoperability between R and Python. Python variables can be accessed from R using py$_var_.

```{r}
library(reticulate)

knitr::kable(py$df)
```

### DataTables

*R*

```{r}
DT::datatable(df)
```

*Python*

```{r}
DT::datatable(py$df)
```
